Thermodynamic response functions in a cell fluid model

### 2024-09-10 paper_text.tex

First, the quantity $M$ depends linearly on the chemical potential
\begin{align}\label{chem_pot}
	&	M = \frac{\tilde\mu}{W(0)} + g_1 - \frac{g_3}{g_4} \tilde D(0) - \frac{1}{6} \frac{g_3^3}{g_4^2}, \\
	&	\tilde\mu=\mu-\mu^*(1+\tau),
\end{align}
where $\mu^*$ is some positive constant, $\tau$ is the relative temperature $\tau = (T - T_c) / T_c$, $T_c$ is the critical temperature. We will call $M$ the effective chemical potential.

@@

The coefficients $g_n$ are given by the formulas:
\begin{align}
	& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  
	\nonumber \\
	& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, 
	\nonumber\\
	& g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, 
	\nonumber\\
	& a_4 = -g_4,
\end{align}
where $T_n(p,\alpha^*)$ are the following special functions
\begin{equation}
	T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n {\rm e}^{-pm^2}.
\end{equation}
Here $\alpha^*=v e^{\beta_c\mu^*}$, and the parameter $p$ has the form
\begin{equation}
	p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
\end{equation}

@@

\subsection{Pressure as a function of temperature and density}
In this work, by density we mean the particle number density $\rho = \langle N \rangle / V$.
In the framework of the grand canonical ensemble the average number of particles $\left\langle N \right\rangle$ is found by 
\begin{eqnarray*}
	\langle N \rangle & = & \left(\frac{\partial \ln\Xi}{\partial \beta \mu}\right)_{T,V} 
	\\
	& = & V \left(\frac{\partial P}{\partial \mu}\right)_{T,V}.
\end{eqnarray*}
From this equation it follows
\begin{equation*}
	\frac{\langle N \rangle}{V} = \frac{\langle N \rangle}{v N_v} = \left(\frac{\partial P}{\partial \mu}\right)_{T,V}
\end{equation*}
and thus
\begin{equation}
	\bar{n} \equiv \frac{\langle N \rangle}{V} v = v \left(\frac{\partial P}{\partial \mu}\right)_{T,V} 
	= \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
The quantity $\bar{n}$, on the one hand, has the meaning of the reduced particle number density, and is usually denoted by $\rho^*$ in the literature on simple fluids~\cite{HansenMcDonald2013}. In the context of the cell model, on the other hand, it is the average number of particles per cell.
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\bar n = n_g - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $n_g$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020}
\begin{eqnarray}\label{eq:crit_dens}
	n_g & = & g_1 - \frac{g_3}{g_4}\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right) - \frac{1}{6} \frac{g_3^3}{g_4^2} 
	\nonumber\\
	& = & g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is
\begin{equation*}
	n_g = 0.978.
\end{equation*}
Equations~\eqref{eq:density} and~\eqref{eq:eosMT} jointly define a parametric relationship between pressure and density, with $M$ serving as the parameter. Figure~\ref{fig1} shows the dependency of the reduced pressure $P^*$ on the density $\bar{n}$ for various values of the reduced temperature $\tau$, at and above the critical temperature.

This dependency can also be expressed explicitly. To achieve this, we combine the equations~\eqref{eq:density} and~\eqref{eq:ro_M} to express the effective chemical potential $M$ as a function of density and temperature:
\begin{equation}\label{eq:M_nT}
	\bar M = \frac{\rho_{n}}{\beta W(0)} - (\bar n - n_g),
\end{equation}
where
\begin{align} \label{eq:ro_nT}
	& \rho_{n} = - 2 \left(\frac{g_3^2 - 2g_2 g_4}{g_4^2} \right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \\
	& \alpha_n = \arccos \left( \left( - \frac{9 g_4^4}{\left( 2 g_2 g_4 - g_3^2\right)^3}\right)^{\frac{1}{2}} (n_g - \bar n)\right). \nonumber 
\end{align}
The notation $\bar{M}$ represents the effective chemical potential as a function of temperature and density $\bar{n}$, while $M$ represents the effective chemical potential as a function of temperature and chemical potential~$\mu$.

At $T>T_c$ the equation of state of a cell fluid model in terms of the density and the temperature has the following form
\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\eta (\bar n,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4.
\end{equation}
The quantity $E_\nu (\bar n, T)$ in \eqref{eq:ro_nT} is the function $E_\mu (M,T)$ \eqref{eq:E_mu} rewritten in terms of density and temperature taking into account the expression \eqref{eq:M_nT}
\begin{eqnarray}\label{eq:E_nu}
	E_\nu (\bar{n}, T) & = & - \frac{\ln (2\pi \beta W(0))}{2 N_v}  +  g_0 
	\nonumber\\
	&& - \frac{\beta W(0)}{2} 
	\left(\bar{M} - g_1 + \frac{g_3}{g_4} \tilde{D}(0) + \frac{g_3^3}{6g_4^2} \right)^{2} 
	\nonumber\\
	&& - \frac{g_3}{g_4} {\bar{M}} - \frac{g_3^2}{2 g_4^2}  \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
\end{eqnarray}
For the reduced pressure we write
\begin{eqnarray}
	\label{eq:eosPTn_reduced}
	P^* & = & (1 + \tau)T^*_c \bigg[ E_\eta (\bar{n},T) 
	\nonumber\\
	&& + \bar{M} \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4\bigg].
\end{eqnarray}

Figure~\ref{fig1} shows the isotherms for the pressure $P^*$ versus density $\bar{n}$ (left-hand side) and versus the effective chemical potential $M$ (right-hand side).
\begin{figure}[h!]
	\centering 
	\includegraphics[width=0.446\textwidth]{f1a1.pdf} 
	\includegraphics[width=0.45\textwidth]{f1b1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the density $\bar n$ (on the left-hand side) and a function of the effective chemical potential $M$ (on the right-hand side) at $T \geq T_c$ represented by black lines. Thick grey lines on both figures correspond to isotherms of pressure at $T < T_c$ based on the results taken from~\cite{KozlovskiiDobush2020}. 
	}\label{fig1}
\end{figure}

Thus, in this Section, we presented two forms of the equation of state. The first one expresses the pressure as a function of temperature and chemical potential, $P = P(T, \mu)$, and is represented by equivalent Eqs.~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}. The second one expresses the pressure as a function of temperature and density, $P = P(T, \bar{n})$, and is represented by equivalent Eqs.~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}. These equations, together with the explicit dependency of density on temperature and chemical potential, $\bar{n} = \bar{n}(T, \mu)$, Eq.~\eqref{eq:density}, are the basis for the calculation of the thermodynamic response functions in the next Section~\ref{sec:respons_functions}.

@@

Let us introduce the reduced isothermal compressibility
\begin{equation}
	K^*_T \equiv \frac{\varepsilon K_T}{v}.
\end{equation}
The quantity $K^*_T$ is dimensionless and is of order unity, except at the critical point itself, where it is divergent. It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
\begin{eqnarray}
	K^*_T & = & \frac{\varepsilon}{\bar{n}^2} \left(\frac{\partial \bar{n}}{\partial \mu}\right)_T
	\nonumber\\
	& = & \frac{\varepsilon^2}{\bar{n}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
Now, either Eq.~\eqref{eq:eosPTM_reduced} or Eq.~\eqref{eq:density} can be used to explicitly calculate $K^*_T$, with the result expressed as a function of temperature and chemical potential.

If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the isothermal compressibility are
\begin{equation}
	K_T = \frac{1}{\bar{n}} \left(\frac{\partial P}{\partial \bar{n}}\right)^{-1}_T
\end{equation}
and 
\begin{equation}
	K^*_T = \frac{1}{\bar{n}} \left(\frac{\partial P^*}{\partial \bar{n}}\right)^{-1}_T.
\end{equation}

Figure~\ref{fig2a} illustrates the dependency of the compressibility $K^*_T$ on the density $\bar{n}$ for various values of temperature above the critical one. The dependency of $K^*_T$ on the effective chemical potential $M$ is displayed in Fig~\ref{fig2b}.

\begin{figure}[htbp]
	%\centering 
	\includegraphics[width=0.5\textwidth]{f2a.pdf}
	%\includegraphics[width=\column]{f2a}
	%\vskip-3mm
	\caption{The reduced isothermal compressibility $K^*_T$ as a function of the density $\bar n$ at different values of reduced temperature $\tau > 0$ ($T > T_c$). 
	}
	\label{fig2a}
\end{figure}

@@

\subsection{Thermal pressure coefficient}
The thermal pressure coefficient is defined by
\begin{equation}
	\label{def:therm_pres_coef}
	\beta_V = \left( \frac{\partial P}{\partial T} \right)_{V,N}.
\end{equation}
We rewrite $\beta_V$ into a form that is suitable for the equation of state $P=P(T,\mu)$, see~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}
\begin{eqnarray}
	\beta_V	& = & \left(\frac{\partial P}{\partial T}\right)_{\mu} + \left(\frac{\partial P}{\partial \mu}\right)_T \left(\frac{\partial \mu}{\partial T}\right)_{V, N}.
\end{eqnarray}
Applying the cyclic relation
\begin{equation*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	\left(\frac{\partial T}{\partial V}\right)_{\mu, N}
	\left(\frac{\partial V}{\partial \mu}\right)_{T, N}
	= -1
\end{equation*}
we obtain
\begin{eqnarray*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	& = & 
	- \left(\frac{\partial V}{\partial T}\right)_{\mu, N} 
	\left(\frac{\partial V}{\partial \mu}\right)^{-1}_{T, N} \\
	& = & - \left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
	\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}.		
\end{eqnarray*}
Substituting this result into the formula for $\beta_V$, we arrive at the final expression for the thermal pressure coefficient
\begin{equation}
	\beta_V = \left(\frac{\partial P}{\partial T}\right)_{\mu} 
	- \left(\frac{\partial P}{\partial \mu}\right)_T 
	\left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
	\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}
\end{equation}
which\footnote{Compare this equation for $\beta_V$ with Eq.(17) from~\cite{StrokerMeier2021}} is easily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}. It is also worth noting that the first contribution to $\beta_V$ is essentially the entropy per volume, $S/V = (\partial P / \partial T)_\mu$.

We introduce the reduced thermal pressure coefficient by
\begin{equation}
	\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
\end{equation}
It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
\begin{equation}
	\label{eq:beta_star_m}
	\beta^*_V = \frac{1}{T^*_c} 
	\left[ \left(\frac{\partial P^*}{\partial \tau}\right)_{\mu} 
	- \left(\frac{\partial P^*}{\partial \mu}\right)_T 
	\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
	\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T} 
	\right].
\end{equation}

If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the thermal pressure coefficient follow immediately from the definition~\eqref{def:therm_pres_coef}
\begin{equation}
	\beta_V = \left(\frac{\partial P}{\partial T}\right)_\rho
\end{equation}
and
\begin{equation}
	\label{eq:beta_star_n}
	\beta^*_V = \frac{1}{T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho}.
\end{equation}

\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f3a.pdf}
	\vskip-3mm\caption{The reduced thermal pressure coefficient $\beta^*_V$ as a function of the density $\bar n$ at different values of relative temperature $\tau > 0$ ($T > T_c$). 
	}\label{fig3a}
\end{figure}
\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f3b.pdf}
	\vskip-3mm\caption{The reduced thermal pressure coefficient $\beta^*_V$ as a function of the effective chemical potential $M$ at different values of relative temperature $\tau > 0$ ($T > T_c$). 
	}\label{fig3b}
\end{figure}
\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f3c.pdf}
	\vskip-3mm\caption{The reduced thermal pressure coefficient $\beta^*_V$ as a function of the relative temperature $\tau$ at different values of the density $\eta$. 
	}\label{fig3c}
\end{figure}

Figure~\ref{fig3a} shows the dependency of the pressure coefficient $\beta^*_V$ on the density $\bar{n}$ for various values of temperature above the critical one. The dependency of $\beta^*_V$ on the effective chemical potential $M$ is displayed in Fig~\ref{fig3b}.

@@

We introduce the reduced thermal expansion coefficient by
\begin{equation}
	\alpha^*_P = \frac{\varepsilon}{k_{\rm B}} \alpha_P.
\end{equation}
It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
\begin{eqnarray}
	\alpha^*_P & = & \frac{1}{T^*_c \bar{n}}
	\left[ 
	-\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
	\right. \nonumber\\
	&& + \left.\left(\frac{\partial \bar{n}}{\partial \mu}\right)_{T}
	\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
	\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
	\right].
\end{eqnarray}

If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then we need to rewrite the definition for $\alpha_P$ accordingly. We start with~\eqref{eq:alpha_vs_rho} and apply the cyclic relation
\begin{equation}
	\left(\frac{\partial \rho}{\partial T}\right)_P = - \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T
\end{equation}
to get for $\alpha_P$
\begin{equation}
	\alpha_P = \frac{1}{\rho} \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T.
\end{equation}
or for $\alpha^*_P$
\begin{equation}
	\alpha^*_P = \frac{1}{\bar{n} T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho} \left(\frac{\partial P^*}{\partial \bar{n}}\right)^{-1}_T.
\end{equation}

Figure~\ref{fig4a} shows the dependency of the reduced thermal expansion coefficient $\alpha^*_P$ on the density $\bar{n}$ for various values of temperature above the critical one. The dependency of $\alpha^*_P$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig4b}.

\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f4a.pdf}
	\vskip-3mm\caption{The reduced thermal expansion coefficient $\alpha^*_P$ as a function of the density $\bar n$ at different values of relative temperature $\tau > 0$ ($T > T_c$). 
	}\label{fig4a}
\end{figure}

>>>>

First, the quantity $M$ depends linearly on the chemical potential
\begin{align}\label{chem_pot}
	&	M = \frac{\tilde\mu}{W(0)} + g_1 - \frac{g_3}{g_4} \tilde D(0) - \frac{1}{6} \frac{g_3^3}{g_4^2}, \\
	&	\tilde\mu=\mu-\mu_0(1+\tau),
\end{align}
where $\mu_0$ is some positive constant, $\tau$ is the relative temperature $\tau = (T - T_c) / T_c$, $T_c$ is the critical temperature. We will call $M$ the effective chemical potential.

@@

The coefficients $g_n$ are given by the formulas:
\begin{align}
	& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  
	\nonumber \\
	& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, 
	\nonumber\\
	& g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, 
	\nonumber\\
	& a_4 = -g_4,
\end{align}
where $T_n(p,\alpha^*)$ are the following special functions
\begin{equation}
	T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n {\rm e}^{-pm^2}.
\end{equation}
Here $\alpha^*=v e^{\beta_c\mu_0}$, and the parameter $p$ has the form
\begin{equation}
	p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
\end{equation}

@@

\subsection{Pressure as a function of temperature and density}
In this work, by density we mean the particle number density $\rho = \langle N \rangle / V$.
In the framework of the grand canonical ensemble the average number of particles $\left\langle N \right\rangle$ is found by 
\begin{eqnarray*}
	\langle N \rangle & = & \left(\frac{\partial \ln\Xi}{\partial \beta \mu}\right)_{T,V} 
	\\
	& = & V \left(\frac{\partial P}{\partial \mu}\right)_{T,V}.
\end{eqnarray*}
From this equation it follows
\begin{equation*}
	\frac{\langle N \rangle}{V} = \frac{\langle N \rangle}{v N_v} = \left(\frac{\partial P}{\partial \mu}\right)_{T,V}
\end{equation*}
and thus
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = v \left(\frac{\partial P}{\partial \mu}\right)_{T,V} 
	= \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
The quantity $\rho^*$, on the one hand, is the reduced particle number density, which is the notation commonly used in the literature on simple fluids~\cite{HansenMcDonald2013}. In the context of the cell model, on the other hand, it is the average number of particles per cell.\footnote{In our previous works we denoted the reduced number density by $\bar{n}$. In the current work we switch to more common notation $\rho^*$.}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = n_g - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $n_g$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020}
\begin{eqnarray}\label{eq:crit_dens}
	n_g & = & g_1 - \frac{g_3}{g_4}\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right) - \frac{1}{6} \frac{g_3^3}{g_4^2} 
	\nonumber\\
	& = & g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is
\begin{equation*}
	n_g = 0.978.
\end{equation*}
Equations~\eqref{eq:density} and~\eqref{eq:eosMT} jointly define a parametric relationship between pressure and density, with $M$ serving as the parameter. Figure~\ref{fig1} shows the dependency of the reduced pressure $P^*$ on the density $\rho^*$ for various values of the reduced temperature $\tau$, at and above the critical temperature.

This dependency can also be expressed explicitly. To achieve this, we combine the equations~\eqref{eq:density} and~\eqref{eq:ro_M} to express the effective chemical potential $M$ as a function of density and temperature:
\begin{equation}\label{eq:M_nT}
	\bar M = \frac{\rho_{n}}{\beta W(0)} - (\rho^* - n_g),
\end{equation}
where
\begin{align} \label{eq:ro_nT}
	& \rho_{n} = - 2 \left(\frac{g_3^2 - 2g_2 g_4}{g_4^2} \right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \\
	& \alpha_n = \arccos \left( \left( - \frac{9 g_4^4}{\left( 2 g_2 g_4 - g_3^2\right)^3}\right)^{\frac{1}{2}} (n_g - \rho^*)\right). \nonumber 
\end{align}
The notation $\bar{M}$ represents the effective chemical potential as a function of temperature and density $\rho^*$, while $M$ represents the effective chemical potential as a function of temperature and chemical potential~$\mu$.

At $T>T_c$ the equation of state of a cell fluid model in terms of the density and the temperature has the following form
\begin{equation}\label{eq:eosNT}
	Pv\beta = E_\eta (\rho^*,T) + \bar M \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4.
\end{equation}
The quantity $E_\nu (\rho^*, T)$ in \eqref{eq:ro_nT} is the function $E_\mu (M,T)$ \eqref{eq:E_mu} rewritten in terms of density and temperature taking into account the expression \eqref{eq:M_nT}
\begin{eqnarray}\label{eq:E_nu}
	E_\nu (\rho^*, T) & = & - \frac{\ln (2\pi \beta W(0))}{2 N_v}  +  g_0 
	\nonumber\\
	&& - \frac{\beta W(0)}{2} 
	\left(\bar{M} - g_1 + \frac{g_3}{g_4} \tilde{D}(0) + \frac{g_3^3}{6g_4^2} \right)^{2} 
	\nonumber\\
	&& - \frac{g_3}{g_4} {\bar{M}} - \frac{g_3^2}{2 g_4^2}  \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
\end{eqnarray}
For the reduced pressure we write
\begin{eqnarray}
	\label{eq:eosPTn_reduced}
	P^* & = & (1 + \tau)T^*_c \bigg[ E_\eta (\rho^*,T) 
	\nonumber\\
	&& + \bar{M} \rho_{n} + \frac{\tilde D(0)}{2} \rho_{n}^2 - \frac{a_4}{24} \rho_{n}^4\bigg].
\end{eqnarray}

Figure~\ref{fig1} shows the isotherms for the pressure $P^*$ versus density $\rho^*$ (left-hand side) and versus the effective chemical potential $M$ (right-hand side).
\begin{figure}[h!]
	\centering 
	\includegraphics[width=0.446\textwidth]{f1a1.pdf} 
	\includegraphics[width=0.45\textwidth]{f1b1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the density $\rho^*$ (on the left-hand side) and a function of the effective chemical potential $M$ (on the right-hand side) at $T \geq T_c$ represented by black lines. Thick grey lines on both figures correspond to isotherms of pressure at $T < T_c$ based on the results taken from~\cite{KozlovskiiDobush2020}. 
	}\label{fig1}
\end{figure}

Thus, in this Section, we presented two forms of the equation of state. The first one expresses the pressure as a function of temperature and chemical potential, $P = P(T, \mu)$, and is represented by equivalent Eqs.~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}. The second one expresses the pressure as a function of temperature and density, $P = P(T, \rho^*)$, and is represented by equivalent Eqs.~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}. These equations, together with the explicit dependency of density on temperature and chemical potential, $\rho^* = \rho^*(T, \mu)$, Eq.~\eqref{eq:density}, are the basis for the calculation of the thermodynamic response functions in the next Section~\ref{sec:respons_functions}.

@@

Let us introduce the reduced isothermal compressibility
\begin{equation}
	K^*_T \equiv \frac{\varepsilon K_T}{v}.
\end{equation}
The quantity $K^*_T$ is dimensionless and is of order unity, except at the critical point itself, where it is divergent. It is expressed in terms of the reduced quantities $P^*$ and $\rho^*$ as follows
\begin{eqnarray}
	K^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
	\nonumber\\
	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
Now, either Eq.~\eqref{eq:eosPTM_reduced} or Eq.~\eqref{eq:density} can be used to explicitly calculate $K^*_T$, with the result expressed as a function of temperature and chemical potential.

If it is preferable to use the equation of state in the form $P=P(T, \rho^*)$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the isothermal compressibility are
\begin{equation}
	K_T = \frac{1}{\rho^*} \left(\frac{\partial P}{\partial \rho^*}\right)^{-1}_T
\end{equation}
and 
\begin{equation}
	K^*_T = \frac{1}{\rho^*} \left(\frac{\partial P^*}{\partial \rho^*}\right)^{-1}_T.
\end{equation}

Figure~\ref{fig2a} illustrates the dependency of the compressibility $K^*_T$ on the density $\rho^*$ for various values of temperature above the critical one. The dependency of $K^*_T$ on the effective chemical potential $M$ is displayed in Fig~\ref{fig2b}.

\begin{figure}[htbp]
	%\centering 
	\includegraphics[width=0.5\textwidth]{f2a.pdf}
	%\includegraphics[width=\column]{f2a}
	%\vskip-3mm
	\caption{The reduced isothermal compressibility $K^*_T$ as a function of the density $\rho^*$ at different values of reduced temperature $\tau > 0$ ($T > T_c$). 
	}
	\label{fig2a}
\end{figure}

@@

\subsection{Thermal pressure coefficient}
The thermal pressure coefficient is defined by
\begin{equation}
	\label{def:therm_pres_coef}
	\beta_V = \left( \frac{\partial P}{\partial T} \right)_{V,N}.
\end{equation}
We rewrite $\beta_V$ into a form that is suitable for the equation of state $P=P(T,\mu)$, see~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}
\begin{eqnarray}
	\beta_V	& = & \left(\frac{\partial P}{\partial T}\right)_{\mu} + \left(\frac{\partial P}{\partial \mu}\right)_T \left(\frac{\partial \mu}{\partial T}\right)_{V, N}.
\end{eqnarray}
Applying the cyclic relation
\begin{equation*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	\left(\frac{\partial T}{\partial V}\right)_{\mu, N}
	\left(\frac{\partial V}{\partial \mu}\right)_{T, N}
	= -1
\end{equation*}
we obtain
\begin{eqnarray*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	& = & 
	- \left(\frac{\partial V}{\partial T}\right)_{\mu, N} 
	\left(\frac{\partial V}{\partial \mu}\right)^{-1}_{T, N} \\
	& = & - \left(\frac{\partial \rho^*}{\partial T}\right)_{\mu}
	\left(\frac{\partial \rho^*}{\partial \mu}\right)^{-1}_{T}.		
\end{eqnarray*}
Substituting this result into the formula for $\beta_V$, we arrive at the final expression for the thermal pressure coefficient
\begin{equation}
	\beta_V = \left(\frac{\partial P}{\partial T}\right)_{\mu} 
	- \left(\frac{\partial P}{\partial \mu}\right)_T 
	\left(\frac{\partial \rho^*}{\partial T}\right)_{\mu}
	\left(\frac{\partial \rho^*}{\partial \mu}\right)^{-1}_{T}
\end{equation}
which\footnote{Compare this equation for $\beta_V$ with Eq.(17) from~\cite{StrokerMeier2021}} is easily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}. It is also worth noting that the first contribution to $\beta_V$ is essentially the entropy per volume, $S/V = (\partial P / \partial T)_\mu$.

We introduce the reduced thermal pressure coefficient by
\begin{equation}
	\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
\end{equation}
It is expressed in terms of reduced quantities as follows
\begin{equation}
	\label{eq:beta_star_m}
	\beta^*_V = \frac{1}{T^*_c} 
	\left[ \left(\frac{\partial P^*}{\partial \tau}\right)_{\mu} 
	- \left(\frac{\partial P^*}{\partial \mu}\right)_T 
	\left(\frac{\partial \rho^*}{\partial \tau}\right)_{\mu}
	\left(\frac{\partial \rho^*}{\partial \mu}\right)^{-1}_{T} 
	\right].
\end{equation}

If it is preferable to use the equation of state in the form $P=P(T, \rho^*)$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the thermal pressure coefficient follow immediately from the definition~\eqref{def:therm_pres_coef}
\begin{equation}
	\beta_V = \left(\frac{\partial P}{\partial T}\right)_\rho
\end{equation}
and
\begin{equation}
	\label{eq:beta_star_n}
	\beta^*_V = \frac{1}{T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho}.
\end{equation}

\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f3a.pdf}
	\vskip-3mm\caption{The reduced thermal pressure coefficient $\beta^*_V$ as a function of the density $\rho^*$ at different values of relative temperature $\tau > 0$ ($T > T_c$). 
	}\label{fig3a}
\end{figure}
\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f3b.pdf}
	\vskip-3mm\caption{The reduced thermal pressure coefficient $\beta^*_V$ as a function of the effective chemical potential $M$ at different values of relative temperature $\tau > 0$ ($T > T_c$). 
	}\label{fig3b}
\end{figure}
\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f3c.pdf}
	\vskip-3mm\caption{The reduced thermal pressure coefficient $\beta^*_V$ as a function of the relative temperature $\tau$ at different values of the density $\rho^*$. 
	}\label{fig3c}
\end{figure}

Figure~\ref{fig3a} shows the dependency of the pressure coefficient $\beta^*_V$ on the density $\rho^*$ for various values of temperature above the critical one. The dependency of $\beta^*_V$ on the effective chemical potential $M$ is displayed in Fig~\ref{fig3b}.

@@

We introduce the reduced thermal expansion coefficient by
\begin{equation}
	\alpha^*_P = \frac{\varepsilon}{k_{\rm B}} \alpha_P.
\end{equation}
It is expressed in terms of reduced quantities as follows
\begin{eqnarray}
	\alpha^*_P & = & \frac{1}{T^*_c \rho^*}
	\left[ 
	-\left(\frac{\partial \rho^*}{\partial \tau}\right)_{\mu}
	\right. \nonumber\\
	&& + \left.\left(\frac{\partial \rho^*}{\partial \mu}\right)_{T}
	\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
	\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
	\right].
\end{eqnarray}

If it is preferable to use the equation of state in the form $P=P(T, \rho^*)$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then we need to rewrite the definition for $\alpha_P$ accordingly. We start with~\eqref{eq:alpha_vs_rho} and apply the cyclic relation
\begin{equation}
	\left(\frac{\partial \rho}{\partial T}\right)_P = - \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T
\end{equation}
to get for $\alpha_P$
\begin{equation}
	\alpha_P = \frac{1}{\rho} \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T.
\end{equation}
or for $\alpha^*_P$
\begin{equation}
	\alpha^*_P = \frac{1}{\rho^* T^*_c} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho} \left(\frac{\partial P^*}{\partial \rho^*}\right)^{-1}_T.
\end{equation}

Figure~\ref{fig4a} shows the dependency of the reduced thermal expansion coefficient $\alpha^*_P$ on the density $\rho^*$ for various values of temperature above the critical one. The dependency of $\alpha^*_P$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig4b}.

\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f4a.pdf}
	\vskip-3mm\caption{The reduced thermal expansion coefficient $\alpha^*_P$ as a function of the density $\rho^*$ at different values of relative temperature $\tau > 0$ ($T > T_c$). 
	}\label{fig4a}
\end{figure}
